---
title: "Hands On Exercise 3.1"
format: html
author: "Bryan Bramaskara"
date: "September 11, 2025"
date-modified: "last-modified"
execute: 
  eval: false
  echo: true
  warning: false
  freeze: true
---

# 6. Spatio-Temporal Point Patterns Analysis

## **6.1 Overview**

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event. Examples of events include incidence of disease, sightings or births of a species, or the occurrences of fires, earthquakes, lightning strikes, tsunamis, or volcanic eruptions.

The analysis of spatio-temporal point patterns is becoming increasingly necessary, given the rapid emergence of geographically and temporally indexed data in a wide range of fields. Several spatio-temporal point patterns analysis methods have been introduced and implemented in R in the last ten years. This chapter shows how various R packages can be combined to run a set of spatio-temporal point pattern analyses in a guided and intuitive way. A real world forest fire events in Kepulauan Bangka Belitung, Indonesia from 1st January 2023 to 31st December 2023 is used to illustrate the methods, procedures and interpretations.

## 6.2 Research Questions

The specific questions we would like to answer are:

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

## 6.3 The data

For the purpose of this exercise, two data sets will be used, they are:

-   *forestfires*, a csv file provides locations of forest fire detected from the Moderate Resolution Imaging Spectroradiometer (MODIS) sensor data. The data are downloaded from [Fire Information for Resource Management System](https://firms.modaps.eosdis.nasa.gov/download/). For the purpose of this exercise, only forest fires within Kepulauan Bangka Belitung will be used.

-   *Kepulauan_Bangka_Belitung*, an ESRI shapefile showing the sub-district (i.e. [kelurahan](https://en.wikipedia.org/wiki/Villages_of_Indonesia)) boundary of Kepulauan Bangka Belitung. The data set was downloaded from [Indonesia Geospatial](https://www.indonesia-geospasial.com/2023/05/download-shapefile-batas-administrasi.html) portal. The original data covers the whole Indonesia. For the purpose of this exercise, only sub-districts within Kepulauan Bangka Belitung are extracted.

## 6.4 Installing and loading R packages

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

## 6.5 Importing and Preparing Study Area

### 6.5.1 Importing study area

```{r}
kbb_sf <- st_read(dsn="data/rawdata",
               layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

### 6.5.2 Converting OWIN

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin

class(kbb_owin)
```

## 6.6 Importing and Preparing Forest Fire data

```{r}
fire_sf <- read_csv("data/rawdata/modis/modis_2024_Indonesia.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
                       crs = 4326) %>%
  st_transform(crs = 32748)
```

Because **ppp** object only accept numerical or character as mark. The code chunk below is used to convert data type of *acq_date* to numeric.

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE))
```

## 6.7 Visualising the Fire Points

### 6.7.1 Overall plot

```{r}
tm_shape(kbb_sf) + 
  tm_polygons()
tm_shape(fire_sf) +
  tm_dots()
```

### 6.7.2 Visualising geographic distribution of forest fires by month

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE)
```

## 6.8. Computing STKDE by month

### 6.8.1 Extracting forest fires by month

```{r}
fire_month <- fire_sf %>% 
  select(Month_num)
```

### 6.8.2 Creating ppp

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp

summary(fire_month_ppp)

any(duplicated(fire_month_ppp))
```

### 6.8.3 Including Owin Object

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)

plot(fire_month_owin)
```

### 6.8.4 Computing Spatio-temporal KDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)

```

### 6.8.5 Plotting the spatio-temporal KDE object

```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

## 6.9 Computing STKDE by Day of Year

### 6.9.1 Creating ppp object

```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayofYear) %>%
  as.ppp()
```

### 6.9.2 Including Owin Object

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

### 6.9.3 Spatio-temporal Density Estimate

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)

plot(kde_yday)
```

## 6.10 Computing STKDE by Day of Year: Improved method

One of the nice function provides in sparr package is `BOOT.spattemp()`. It support bandwidth selection for standalone spatiotemporal density/intensity based on bootstrap estimation of the MISE, providing an isotropic scalar spatial bandwidth and a scalar temporal bandwidth.

Code chunk below uses `BOOT.spattemp()` to determine both the spatial bandwidth and the scalar temporal bandwidth.

```{r}
set.seed(1234)
BOOT.spattemp(fire_yday_owin) 
```

### 6.10.1 Computing spatio-temporal KDE

```{r}
kde_yday <- spattemp.density(fire_yday_owin,
                             h = 9000,
                             lambda = 19)
summary(kde_yday)
```

### 6.10.2 Plotting the output spatio-temporal KDE

```{r}
plot(kde_yday)
```
